Towards Verifying the Bitcoin-S Library

### 2019-07-04 paper/paper.tex
more work

\begin{abstract}
  We try to verify some properties of the bitcoin-s library, a Scala
  implementation of parts of the Bitcoin protocol. We use the
  Stainless verifier which supports programs in a subset of Scala
  called \emph{Pure Scala}.  Since bitcoin-s is not written in this
  fragment, we extract the relevant code from it and perform a series
  of equivalent transformations until we arrive at code that we
  successfully verify. In that process we find and fix two bugs in
  bitcoin-s.

@@

\paragraph{Extracting the Relevant Code} The relevant code for the
addition of satoshis is in two files: CurrencyUnits.scala and
NumberType.scala. From those files we removed the majority of the code
because it is not needed for the verification of our property.  For
example, we removed all number types except for Int64 (so Int32,
UInt64, etc.) because they are not used. We also removed the
superclasses Factory and NetworkElement of CurrencyUnit and Number,
respectively, because the inherited members are not used. Also, we
removed all binary operations on Number that are not used, like
subtraction and multiplication. The extracted code is shown in
Figure~\ref{fig:numbertype} and Figure~\ref{fig:currencyunits}.

@@

We have show
-- bugs (already fixed)
-- suggestions for bitcoin-s

Because of the limitations of the verication tool, we could only
verify a rewritten version of the original Bitcoin-S code.

So code should be written specically with formal verication in mind,

Also, we found that trying to verify code reveals bugs as shown in
section \ref{sec:bugfix}.  Finally, our work led to some feedback to
the Stainless developers to improve the tool.

@@

>>>>

\begin{abstract}
  We try to verify some properties of the bitcoin-s library, a Scala
  implementation of parts of the Bitcoin protocol. We use the
  Stainless verifier which supports programs in a fragment of Scala
  called \emph{Pure Scala}.  Since bitcoin-s is not written in this
  fragment, we extract the relevant code from it and perform a series
  of equivalent transformations until we arrive at code that we
  successfully verify. In that process we find and fix two bugs in
  bitcoin-s.

@@

\paragraph{Extracting the Relevant Code} The relevant code for the
addition of satoshis is in two files: CurrencyUnits.scala and
NumberType.scala. From those files we removed the majority of the code
because it is not needed for the verification of our property.  For
example, we removed all number types except for Int64 (so Int32,
UInt64, etc.) because they are not used. We also removed the
superclasses Factory and NetworkElement of CurrencyUnit and Number,
respectively, because the inherited members are not used. We further
removed all binary operations on Number that are not used, like
subtraction and multiplication. The extracted code is shown in
Figure~\ref{fig:numbertype} and Figure~\ref{fig:currencyunits}.

@@

We are happy to see some friendly green verifier output. However,
apart from the bugs we found, the main conclusion from this work is
that we had to non-trivially transform even a very small portion of
the code in order to verify it. At the moment, it is unrealistic to
routinely formally verify properties as part of the bitcoin-s
development process. However, Stainless development has already
progressed (e.g.\ type members are supported in recent versions) and
continues to do so (e.g. self-referential type bounds are on the
roadmap). Some missing features that we identified are presumably very
easy to support, like the message parameter in the require
function. Some other features presumably require more substantial
work, like bitwise operations on integer types.

On the other hand, bitcoin-s uses features that might not be supported
even by future Stainless versions, such as (concrete, non-case)
classes \todo{is that even true? where?} or calls to Java code. Here
the bitcoin-s code can hopefully be adapted to accomodate formal
verification.

From our results we conclude that formal verification of bitcoin
libraries in general and bitcoin-s in particular is a worthwile
endeavour. We are looking forward to verifying more substantial parts
of the code in future work.

@@

\todo{reference to verified code}
